---
title: "Section 3: Random Data Generation for MaxDiff and Conjoint Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 3. Random Data Generation

## Simulate Conjoint choice tasks

```{r conjoint-simulation}
library(tidyverse)

# Define attributes and levels
smartphone_attrs <- list(
  brand = c("Apple", "Samsung", "Google", "Xiaomi"),
  price = c("$400", "$700", "$1000"),
  battery = c("15 hours", "20 hours", "25 hours"),
  camera = c("16 MP", "64 MP", "128 MP")
)

# Function to generate a random profile
generate_profile <- function(attrs) {
  profile <- list()
  for (attr_name in names(attrs)) {
    profile[[attr_name]] <- sample(attrs[[attr_name]], 1)
  }
  return(profile)
}

# Function to generate a choice set with n alternatives
generate_choice_set <- function(attrs, n_alts = 3) {
  choice_set <- list()
  for (i in 1:n_alts) {
    choice_set[[i]] <- generate_profile(attrs)
  }
  return(choice_set)
}

# Generate 5 example choice sets
set.seed(123) # For reproducibility
example_choice_sets <- list()
for (i in 1:5) {
  example_choice_sets[[i]] <- generate_choice_set(smartphone_attrs)
}

# Display the first choice set
choice_set_df <- bind_rows(lapply(example_choice_sets[[1]], as.data.frame))
knitr::kable(choice_set_df, caption = "Example Choice Set for Conjoint Analysis")
```

::: {.interpretation}
The table above shows an example choice set for a conjoint analysis task. Each row represents a different product alternative with varying levels for each attribute. In a real survey, respondents would be asked to choose their preferred option from these alternatives.
:::

```{r conjoint-simulation-with-preferences}
# Simulate respondent choices based on predefined preferences
# Define some preference weights
preference_weights <- list(
  brand = c(Apple = 0.8, Samsung = 0.5, Google = 0.3, Xiaomi = 0.1),
  price = c("$400" = 0.7, "$700" = 0.4, "$1000" = 0.1),
  battery = c("15 hours" = 0.2, "20 hours" = 0.5, "25 hours" = 0.8),
  camera = c("16 MP" = 0.3, "64 MP" = 0.6, "128 MP" = 0.9)
)

# Function to calculate utility for a profile
calculate_profile_utility <- function(profile, weights) {
  utility <- 0
  for (attr_name in names(profile)) {
    level <- profile[[attr_name]]
    utility <- utility + weights[[attr_name]][level]
  }
  return(utility)
}

# Function to simulate choice based on utilities
simulate_choice <- function(choice_set, weights) {
  utilities <- sapply(choice_set, calculate_profile_utility, weights = weights)
  # Add some random noise
  utilities <- utilities + rnorm(length(utilities), 0, 0.2)
  chosen <- which.max(utilities)
  return(chosen)
}

# Simulate choices for the example choice sets
simulated_choices <- sapply(example_choice_sets, simulate_choice, weights = preference_weights)

# Display results
cat("Simulated choices for 5 choice sets:\n")
for (i in 1:length(simulated_choices)) {
  cat(paste0("Choice set ", i, ": Alternative ", simulated_choices[i], " chosen\n"))
}
```

## Simulate MaxDiff tasks

```{r maxdiff-simulation}
# Define the full list of items
features <- c(
  "Long battery life",
  "Fast charging",
  "High-resolution display",
  "Water resistance",
  "Compact design",
  "Expandable storage",
  "Wireless charging",
  "Advanced camera features",
  "Biometric security",
  "5G connectivity"
)

# Function to generate a MaxDiff question set
generate_maxdiff_set <- function(items, items_per_set = 4) {
  sample(items, items_per_set)
}

# Generate 10 MaxDiff question sets
set.seed(456) # For reproducibility
maxdiff_sets <- list()
for (i in 1:10) {
  maxdiff_sets[[i]] <- generate_maxdiff_set(features)
}

# Display the first three MaxDiff sets
maxdiff_example <- data.frame(
  Set = rep(1:3, each = 4),
  Item = unlist(maxdiff_sets[1:3])
)

knitr::kable(maxdiff_example, caption = "Example MaxDiff Question Sets")
```

::: {.interpretation}
The table above shows three example MaxDiff question sets. Each set contains 4 randomly selected items from our list of 10 smartphone features. In a real survey, respondents would select the most and least important feature from each set.
:::

```{r maxdiff-simulation-with-preferences}
# Define true preference scores for the features
true_preferences <- c(
  "Long battery life" = 9,
  "Fast charging" = 7,
  "High-resolution display" = 6,
  "Water resistance" = 4,
  "Compact design" = 3,
  "Expandable storage" = 2,
  "Wireless charging" = 5,
  "Advanced camera features" = 8,
  "Biometric security" = 5,
  "5G connectivity" = 6
)

# Function to simulate best-worst choices based on preferences
simulate_maxdiff_choices <- function(item_set, preferences) {
  # Get preference scores for items in this set
  set_preferences <- preferences[item_set]
  
  # Add some random noise
  set_preferences <- set_preferences + rnorm(length(set_preferences), 0, 0.5)
  
  # Select best and worst
  best_idx <- which.max(set_preferences)
  worst_idx <- which.min(set_preferences)
  
  return(list(
    best = item_set[best_idx],
    worst = item_set[worst_idx]
  ))
}

# Simulate responses for the first three sets
set.seed(789)
maxdiff_responses <- data.frame(
  Set = 1:3,
  stringsAsFactors = FALSE
)

for (i in 1:3) {
  choices <- simulate_maxdiff_choices(maxdiff_sets[[i]], true_preferences)
  maxdiff_responses$Best_Item[i] <- choices$best
  maxdiff_responses$Worst_Item[i] <- choices$worst
}

knitr::kable(maxdiff_responses, caption = "Example MaxDiff Responses")
```
