---
title: "Section 8: Combined Usage of MaxDiff and Conjoint Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

# 7. Combined Usage

## How MaxDiff and Conjoint can complement each other

MaxDiff and Conjoint analysis can be used together in a complementary research program:

1. **Sequential approach**: Use MaxDiff first to identify the most important features, then include only those in a conjoint study
2. **Parallel approach**: Run both studies simultaneously to get different perspectives on preferences
3. **Integrated approach**: Use MaxDiff results to inform the design of conjoint attributes and levels

```{r combined-usage, eval=TRUE}
# Example of using MaxDiff results to inform Conjoint design

# Step 1: Identify top features from MaxDiff
# Create simulated MaxDiff results
set.seed(123)
maxdiff_results <- data.frame(
  item = c(
    "Long battery life",
    "Fast charging",
    "High-resolution display",
    "Water resistance",
    "Compact design",
    "Expandable storage",
    "Wireless charging",
    "Advanced camera features",
    "Biometric security",
    "5G connectivity"
  ),
  score = c(85, 72, 65, 45, 30, 25, 55, 78, 50, 60)
)

# Sort by score
maxdiff_results <- maxdiff_results %>%
  arrange(desc(score))

# Get top 5 features
top_features <- maxdiff_results %>%
  slice_head(n = 5) %>%
  pull(item)

cat("Top 5 features from MaxDiff analysis:\n")
for (i in 1:length(top_features)) {
  cat(paste0(i, ". ", top_features[i], "\n"))
}

# Step 2: Design conjoint study based on top features
# (This would be implemented in an actual survey)
conjoint_attributes_from_maxdiff <- list(
  feature1 = c("Basic", "Advanced"),
  feature2 = c("Basic", "Advanced"),
  feature3 = c("Basic", "Advanced"),
  feature4 = c("Basic", "Advanced"),
  feature5 = c("Basic", "Advanced"),
  price = c("Low", "Medium", "High")
)

cat("\nConjoint attributes based on MaxDiff results:\n")
for (attr in names(conjoint_attributes_from_maxdiff)) {
  cat(paste0(attr, ": ", paste(conjoint_attributes_from_maxdiff[[attr]], collapse = ", "), "\n"))
}
```

::: {.tip}
**Sequential Approach Benefits**

Using MaxDiff before Conjoint:
1. Reduces the number of attributes in the conjoint study, making it more manageable
2. Ensures you focus on attributes that matter most to consumers
3. Improves the statistical efficiency of the conjoint design
4. Reduces respondent fatigue by eliminating less important attributes
5. Provides complementary insights: importance (MaxDiff) and trade-offs (Conjoint)
:::

## When to use both in a research program

| Research Goal | Approach |
|---------------|----------|
| **New Product Development** | Use MaxDiff to identify key features, then Conjoint to optimize configurations |
| **Pricing Strategy** | Use MaxDiff to identify value drivers, then Conjoint to determine price sensitivity |
| **Market Segmentation** | Use both methods to identify different preference patterns across segments |
| **Competitive Analysis** | Use MaxDiff for brand attribute importance, Conjoint for competitive positioning |

```{r combined-research-program, eval=TRUE}
# Example of a combined research program for smartphone development

# Phase 1: MaxDiff to identify important features
cat("Phase 1: MaxDiff Study\n")
cat("Purpose: Identify most important smartphone features\n")
cat("Output: Prioritized list of features\n\n")

# Show MaxDiff results (using our simulated data)
knitr::kable(head(maxdiff_results, 5), 
             caption = "Top 5 Features from MaxDiff")

# Phase 2: Conjoint to optimize product configuration
cat("\nPhase 2: Conjoint Study\n")
cat("Purpose: Determine optimal feature levels and price points\n")
cat("Output: Part-worth utilities and market simulations\n\n")

# Create simulated conjoint results
conjoint_results <- data.frame(
  profile = c("Premium", "Mid-range", "Budget", "Value"),
  share = c(35, 30, 20, 15)
)

# Show Conjoint results
knitr::kable(conjoint_results, 
             caption = "Market Simulation Results from Conjoint")

# Phase 3: Integrated Analysis
cat("\nPhase 3: Integrated Analysis\n")
cat("Purpose: Combine insights from both studies\n")
cat("Output: Optimized product configurations for different segments\n")
```

## Storytelling with both together

```{r storytelling, eval=TRUE}
# Example of how to combine insights from both methods
# This would typically be done with real research data

# Step 1: Identify key features from MaxDiff
# (Using our simulated results)
key_features <- maxdiff_results %>%
  slice_head(n = 3) %>%
  pull(item)

# Step 2: Extract insights about these features from Conjoint
# (Using simulated results)
feature_insights <- data.frame(
  feature = c("Long battery life", "Advanced camera features", "Fast charging"),
  utility = c(0.8, 0.7, 0.5),
  price_sensitivity = c("Medium", "High", "Low")
)

# Step 3: Combine into actionable insights
cat("Combined Insights for Product Development:\n\n")

cat("1. Battery Life is a top priority for consumers (MaxDiff rank: 1)\n")
cat("   - Provides significant utility (0.8) in conjoint trade-offs\n")
cat("   - Medium price sensitivity suggests consumers will pay more for better battery\n")
cat("   - Consider making this a key selling point in marketing\n\n")

cat("2. Camera Quality is highly valued (MaxDiff rank: 2)\n")
cat("   - Strong utility (0.7) but high price sensitivity\n")
cat("   - Worth the investment in higher-quality camera components\n")
cat("   - May need tiered options to address different price points\n\n")

cat("3. Fast Charging is important but less critical (MaxDiff rank: 3)\n")
cat("   - Moderate utility (0.5) with low price sensitivity\n")
cat("   - Could be a standard feature across all models\n")
cat("   - Good secondary selling point in marketing\n\n")

cat("Recommended Product Configuration:\n")
cat("- Focus on extended battery life as primary feature\n")
cat("- Offer tiered camera options to address price sensitivity\n")
cat("- Include fast charging as standard across all models\n")
cat("- Target the 'Tech Enthusiast' segment that values these features most\n")
```

::: {.interpretation}
**Storytelling with Combined Insights**

The example above demonstrates how to weave together insights from both MaxDiff and Conjoint analysis to create a compelling story for product development:

1. MaxDiff identifies which features matter most to consumers
2. Conjoint reveals how consumers make trade-offs between these features
3. Combined, they provide a complete picture of consumer preferences
4. This integrated approach leads to more informed product decisions

The narrative connects the dots between importance (MaxDiff) and utility (Conjoint) to create actionable recommendations.
:::
