---
title: "Section 4: Data Structure & Preparation for MaxDiff and Conjoint Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 4. Data Structure & Preparation

## Structure of Conjoint data

Conjoint data typically comes in one of two formats:

1. **Long format**: Each row represents one alternative in a choice set
2. **Wide format**: Each row represents one choice task with multiple alternatives

```{r conjoint-data-structure}
# Example of conjoint data in long format
set.seed(123)
conjoint_data_long <- tibble(
  resp_id = rep(1:3, each = 12),
  task_id = rep(rep(1:4, each = 3), 3),
  alt_id = rep(1:3, 12),
  choice = rep(0, 36),
  brand = sample(c("Apple", "Samsung", "Google", "Xiaomi"), 36, replace = TRUE),
  price = sample(c("$400", "$700", "$1000"), 36, replace = TRUE),
  battery = sample(c("15 hours", "20 hours", "25 hours"), 36, replace = TRUE),
  camera = sample(c("16 MP", "64 MP", "128 MP"), 36, replace = TRUE)
)

# Simulate choices (one choice per task)
for (r in unique(conjoint_data_long$resp_id)) {
  for (t in unique(conjoint_data_long$task_id)) {
    idx <- which(conjoint_data_long$resp_id == r & conjoint_data_long$task_id == t)
    chosen <- sample(idx, 1)
    conjoint_data_long$choice[chosen] <- 1
  }
}

# Show example of the data
knitr::kable(head(conjoint_data_long), caption = "Conjoint Data in Long Format")
```

::: {.interpretation}
The table above shows conjoint data in long format. Each row represents one alternative in a choice set:
- `resp_id`: Respondent identifier
- `task_id`: Task/question identifier
- `alt_id`: Alternative identifier within a task
- `choice`: 1 if this alternative was chosen, 0 otherwise
- Attribute columns: The levels of each attribute for this alternative
:::

```{r conjoint-data-wide}
# Example of conjoint data in wide format
conjoint_data_wide <- tibble(
  resp_id = rep(1:2, each = 2),
  task_id = rep(1:2, 2),
  choice = rep(1:3, length.out = 4),  # Which alternative was chosen
  
  # Alternative 1
  alt1_brand = sample(c("Apple", "Samsung", "Google", "Xiaomi"), 4, replace = TRUE),
  alt1_price = sample(c("$400", "$700", "$1000"), 4, replace = TRUE),
  alt1_battery = sample(c("15 hours", "20 hours", "25 hours"), 4, replace = TRUE),
  alt1_camera = sample(c("16 MP", "64 MP", "128 MP"), 4, replace = TRUE),
  
  # Alternative 2
  alt2_brand = sample(c("Apple", "Samsung", "Google", "Xiaomi"), 4, replace = TRUE),
  alt2_price = sample(c("$400", "$700", "$1000"), 4, replace = TRUE),
  alt2_battery = sample(c("15 hours", "20 hours", "25 hours"), 4, replace = TRUE),
  alt2_camera = sample(c("16 MP", "64 MP", "128 MP"), 4, replace = TRUE),
  
  # Alternative 3
  alt3_brand = sample(c("Apple", "Samsung", "Google", "Xiaomi"), 4, replace = TRUE),
  alt3_price = sample(c("$400", "$700", "$1000"), 4, replace = TRUE),
  alt3_battery = sample(c("15 hours", "20 hours", "25 hours"), 4, replace = TRUE),
  alt3_camera = sample(c("16 MP", "64 MP", "128 MP"), 4, replace = TRUE)
)

# Show example of the data
knitr::kable(conjoint_data_wide, caption = "Conjoint Data in Wide Format")
```

## Structure of MaxDiff data

MaxDiff data can also be structured in different ways:

1. **Stacked format**: Each row represents one item in a MaxDiff set
2. **Binary format**: Each row represents one MaxDiff question with binary indicators for best/worst

```{r maxdiff-data-structure}
# Define the features again for this section
features <- c(
  "Long battery life",
  "Fast charging",
  "High-resolution display",
  "Water resistance",
  "Compact design",
  "Expandable storage",
  "Wireless charging",
  "Advanced camera features",
  "Biometric security",
  "5G connectivity"
)

# Example of MaxDiff data in stacked format
set.seed(456)
maxdiff_data_stacked <- tibble(
  resp_id = rep(1:3, each = 16),
  set_id = rep(rep(1:4, each = 4), 3),
  item_id = rep(1:10, 5)[1:48],
  item = features[rep(1:10, 5)[1:48]],
  is_best = 0,
  is_worst = 0
)

# Simulate best and worst choices
for (r in unique(maxdiff_data_stacked$resp_id)) {
  for (s in unique(maxdiff_data_stacked$set_id)) {
    idx <- which(maxdiff_data_stacked$resp_id == r & maxdiff_data_stacked$set_id == s)
    best_idx <- sample(idx, 1)
    worst_idx <- sample(idx[idx != best_idx], 1)
    
    maxdiff_data_stacked$is_best[best_idx] <- 1
    maxdiff_data_stacked$is_worst[worst_idx] <- 1
  }
}

# Show example of the data
knitr::kable(head(maxdiff_data_stacked), caption = "MaxDiff Data in Stacked Format")
```

::: {.interpretation}
The table above shows MaxDiff data in stacked format. Each row represents one item in a MaxDiff question set:
- `resp_id`: Respondent identifier
- `set_id`: Set/question identifier
- `item_id`: Item identifier
- `item`: The actual item text
- `is_best`: 1 if this item was selected as best, 0 otherwise
- `is_worst`: 1 if this item was selected as worst, 0 otherwise
:::

```{r maxdiff-data-binary}
# Example of MaxDiff data in binary format
maxdiff_data_binary <- tibble(
  resp_id = rep(1:2, each = 3),
  set_id = rep(1:3, 2),
  items = list(
    c("Long battery life", "Fast charging", "High-resolution display", "Water resistance"),
    c("Compact design", "Expandable storage", "Wireless charging", "Advanced camera features"),
    c("Biometric security", "5G connectivity", "Long battery life", "Fast charging"),
    c("Long battery life", "Fast charging", "High-resolution display", "Water resistance"),
    c("Compact design", "Expandable storage", "Wireless charging", "Advanced camera features"),
    c("Biometric security", "5G connectivity", "Long battery life", "Fast charging")
  ),
  best_item = c(
    "Long battery life", "Wireless charging", "5G connectivity",
    "Fast charging", "Advanced camera features", "Long battery life"
  ),
  worst_item = c(
    "Water resistance", "Compact design", "Biometric security",
    "High-resolution display", "Expandable storage", "5G connectivity"
  )
)

# Show example of the data
knitr::kable(maxdiff_data_binary[1:3,], caption = "MaxDiff Data in Binary Format")
```

## How to reshape survey output for modeling

```{r data-reshaping, eval=TRUE}
# Reshaping conjoint data from wide to long format
conjoint_wide_to_long <- function(data_wide) {
  # This is a simplified example
  data_long <- data_wide %>%
    pivot_longer(
      cols = starts_with("alt"),
      names_to = c("alt_id", ".value"),
      names_pattern = "alt(\\d+)_(.*)"
    )
  
  return(data_long)
}

# Example of reshaping wide to long
conjoint_reshaped <- conjoint_wide_to_long(conjoint_data_wide)
knitr::kable(head(conjoint_reshaped), caption = "Reshaped Conjoint Data")

# Reshaping MaxDiff data for analysis
maxdiff_to_analysis_format <- function(data) {
  # Convert to format suitable for analysis
  data_analysis <- data %>%
    mutate(
      choice = case_when(
        is_best == 1 ~ 1,
        is_worst == 1 ~ -1,
        TRUE ~ 0
      )
    ) %>%
    select(resp_id, set_id, item_id, item, choice)
  
  return(data_analysis)
}

# Example of reshaping MaxDiff data
maxdiff_reshaped <- maxdiff_to_analysis_format(maxdiff_data_stacked)
knitr::kable(head(maxdiff_reshaped), caption = "Reshaped MaxDiff Data")
```

## Handle missing or invalid data

```{r missing-data, eval=TRUE}
# Function to check for and handle missing data in conjoint
check_conjoint_data <- function(data) {
  # Check for missing choices
  missing_choices <- data %>%
    group_by(resp_id, task_id) %>%
    summarize(sum_choices = sum(choice), .groups = "drop") %>%
    filter(sum_choices != 1)
  
  if (nrow(missing_choices) > 0) {
    warning(paste("Found", nrow(missing_choices), "tasks with invalid choice patterns"))
  }
  
  # Check for missing attribute values
  missing_attrs <- data %>%
    summarize(across(c(brand, price, battery, camera), ~sum(is.na(.))))
  
  return(list(missing_choices = missing_choices, missing_attrs = missing_attrs))
}

# Introduce some missing data for demonstration
conjoint_data_with_missing <- conjoint_data_long
conjoint_data_with_missing$choice[1:3] <- 0  # No choice in first task
conjoint_data_with_missing$brand[5] <- NA    # Missing brand value

# Check for issues
missing_check <- check_conjoint_data(conjoint_data_with_missing)
cat("Tasks with invalid choice patterns:\n")
print(missing_check$missing_choices)
cat("\nMissing attribute values:\n")
print(missing_check$missing_attrs)

# Function to check for and handle missing data in MaxDiff
check_maxdiff_data <- function(data) {
  # Check for missing best/worst selections
  missing_selections <- data %>%
    group_by(resp_id, set_id) %>%
    summarize(
      sum_best = sum(is_best),
      sum_worst = sum(is_worst),
      .groups = "drop"
    ) %>%
    filter(sum_best != 1 | sum_worst != 1)
  
  if (nrow(missing_selections) > 0) {
    warning(paste("Found", nrow(missing_selections), "sets with invalid selection patterns"))
  }
  
  return(missing_selections)
}

# Introduce some missing data for demonstration
maxdiff_data_with_missing <- maxdiff_data_stacked
maxdiff_data_with_missing$is_best[1:4] <- 0  # No best choice in first set
maxdiff_data_with_missing$is_worst[9:12] <- 0  # No worst choice in third set

# Check for issues
missing_maxdiff_check <- check_maxdiff_data(maxdiff_data_with_missing)
cat("Sets with invalid selection patterns:\n")
print(missing_maxdiff_check)
```

::: {.warning}
**Handling Missing Data**

For conjoint analysis:
- Tasks with no chosen alternative or multiple chosen alternatives should be flagged
- Missing attribute values can be imputed or the entire alternative can be excluded
- Respondents with excessive missing data may need to be removed

For MaxDiff analysis:
- Sets with no best or worst selection should be flagged
- Sets with multiple best or worst selections should be corrected
- Respondents with excessive missing data may need to be removed
:::
