---
title: "Section 10: Resources & References"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 9. Resources & References

## R Packages for Conjoint and MaxDiff Analysis

### Conjoint Analysis Packages
- **mlogit**: Multinomial logit models for choice data
- **conjoint**: Traditional conjoint analysis
- **support.CEs**: Tools for choice experiments
- **bayesm**: Bayesian inference for marketing problems
- **ChoiceModelR**: Hierarchical Bayes for choice-based conjoint

### MaxDiff Analysis Packages
- **support.BWS**: Best-worst scaling analysis
- **logitr**: Logit models for choice data
- **RSGHB**: Hierarchical Bayes estimation for MaxDiff

```{r resources, eval=TRUE}
# Example of available packages
cat("Conjoint Analysis Packages:\n")
conjoint_packages <- c("mlogit", "conjoint", "support.CEs", "bayesm", "ChoiceModelR")
for (pkg in conjoint_packages) {
  cat(paste0("- ", pkg, "\n"))
}

cat("\nMaxDiff Analysis Packages:\n")
maxdiff_packages <- c("support.BWS", "logitr", "RSGHB")
for (pkg in maxdiff_packages) {
  cat(paste0("- ", pkg, "\n"))
}

# Example code for installing packages
cat("\nExample code for installing packages:\n")
cat("```r\n")
cat("# Install required packages\n")
cat("install.packages(c('mlogit', 'support.CEs'))\n")
cat("\n")
cat("# Load packages\n")
cat("library(mlogit)\n")
cat("library(support.CEs)\n")
cat("```\n")
```

## Additional Resources

### Books and Articles
- Orme, B. (2020). Getting Started with Conjoint Analysis: Strategies for Product Design and Pricing Research. Research Publishers LLC.
- Louviere, J. J., Flynn, T. N., & Marley, A. A. J. (2015). Best-Worst Scaling: Theory, Methods and Applications. Cambridge University Press.

### Online Resources
- Sawtooth Software Documentation: [https://sawtoothsoftware.com/resources/technical-papers](https://sawtoothsoftware.com/resources/technical-papers)
- Displayr MaxDiff Resources: [https://docs.displayr.com/wiki/MaxDiff](https://docs.displayr.com/wiki/MaxDiff)
- Counts Analysis of MaxDiff Data: [https://docs.displayr.com/wiki/Counts_Analysis_of_MaxDiff_Data](https://docs.displayr.com/wiki/Counts_Analysis_of_MaxDiff_Data)
- Analyzing MaxDiff Using Standard Logit Models Using R: [https://docs.displayr.com/wiki/Analyzing_MaxDiff_Using_Standard_Logit_Models_Using_R](https://docs.displayr.com/wiki/Analyzing_MaxDiff_Using_Standard_Logit_Models_Using_R)

## Acknowledgments

This training module was developed to provide a comprehensive introduction to both MaxDiff and Conjoint analysis techniques. The code examples are designed to be educational and may need adaptation for specific research applications.

```{r session-info}
# Session information for reproducibility
sessionInfo()
```
